Popularity Adjusted Block Models are Generalized Random Dot Product Graphs

### 2022-01-11 referee-response.Rmd

10. p8, line 14-21. I don't see why the permutation given by $\Pi$ has $K$ fixed points.  
\textcolor{blue}{
We thank the referee for the comment. 
The purpose of Example 1 is to illustrate this explicitly. 
Comparing matrices $X$ and $Y$, we can see that they share three columns: the first, fifth, and ninth. 
The other columns are permuted. 
From Example 1 (as well as the proof of Theorem 1 for $K = 2$), it should be clear how this property holds for larger $K$. 
The proof of Theorem 1 for arbitrary $K$ should then confirm the intuition based on $K = 2$ and $3$. 
}

11. Section 3.1. I found it confusing that community detection algorithms are detailed when $P$ is assumed to be known and then methods for estimating $P$ are recalled. A sentence in the beginning of this section to expose what is presented could help the reader.

12. p10, line 12. $c^{(i)}$ is not defined.  
\textcolor{blue}{
We thank the referee for pointing this out. 
$c^{(i)}$ is the $i^{th}$ entry of column vector $c$. 
This has been added to the paper.
}

13. p10. I didn't understand the exposition of the SSC algorithm.  
\textcolor{blue}{
We thank the referee for the question. 
We have added more detail to our description of the SSC algorithm. 
In particular, we talk a bit about the intuition/motivation behind the algorithm. 
This isn't added to the description, but $c$ and $c_i$ are column vectors of length $n$ and $c^{(i)}$ and $c_i^{(j)}$ are the $i^{th}$ and $j^{th}$ entries of vectors $c$ and $c_i$ respectively. 
}  
\textcolor{blue}{The description of SSC now reads:
}  
\textcolor{blue}{
The SSC algorithm can be described as follows: 
Given \(X \in \mathbb{R}^{n \times d}\) with vectors
\(x_i^\top \in \mathbb{R}^d\) as rows of \(X\), the optimization problem
\(c_i = \arg\min_{c} \|c\|_1\) subject to \(x_i = X^\top c\) and
\(c^{(i)} = 0\), where $c^{(i)}$ is the $i^{th}$ entry of $c$, is solved for each \(i = [n]\). 
The solutions are collected collected into matrix
\(C = \bigl[ c_1 \mid \cdots \mid c_n \bigr]^\top\) to
construct an affinity matrix \(B = |C| + |C^\top|\). If each \(x_i\) lies
exactly on one of \(K\) subspaces, \(B\) describes an undirected graph
consisting of {\em at least} \(K\) disjoint subgraphs, i.e., \(B_{ij} = 0\) if \(x_i, x_j\) lie on different subspaces. 
The intuition here is that points that lie on the same subspace can be described as linear combinations of each other. 
Thus, for each $c_i$, $c_i^{(j)}$ is zero if $x_i$ and $x_j$ belong to different subspaces and may be nonzero if they do. 
If \(X\) instead represents points near \(K\) subspaces with some noise, 
then this property may only hold approximately and thus 
a final graph partitioning step may be required 
(e.g., edge thresholding or spectral clustering).
}

@@

17. Sections 4 and 5. Why do you use either the ARI criterion or the misclassification rate to assess the recovery of the nodes clustering?  
\textcolor{blue}{
We thank the referee for the question.
The first to datasets in section 5 refer to datasets analyzed by other PABM papers. 
For the Leeds Butterfly dataset, while the paper with which we are comparing \citep{noroozi2019estimation} provides enough details to transform and preprocess the data to match their analysis, they do not provide code to run their clustering algorithm. 
Therefore, we use their reported benchmark value, which is in the form of ARI. 
In order to make an apples-to-apples comparison, we report the ARI for OSC and SSC-ASE here as well. 
For all other analyses (simulation and real data), we use misclassification error (whether rate or count). 
}

18. Do you consider sparsity in the simulated adjacency matrices?

19. Why do you limit your simulation to assortative structure? 
Is it possible to see the results with a larger value of $K$? 
To what extent does the value of $K$ impact the computational burden for the algorithms?  
\textcolor{blue}{
We thank the referee for the questions. 
Regarding disassortativity, as mentioned in previous comments, we have added a third simulation study to address this. 
Performance of the algorithms tested is similar to the other two simulation studies and consistent with the theory. 
Regarding $K$, our algorithms require each $n_k$ to be large enough to span its subspace of dimension $K$, so we focused mostly on smaller values of $K$. 
Cursory numerical experiments suggest that OSC and SSC-ASE behave similarly for larger $K$ provided that $n$ is large enough. 
}  
\textcolor{blue}{
Regarding computational complexity, because both OSC and SSC-ASE may have an arbitrary final processing step, we will focus on the complexity in constructing the affinity matrix $B$. 
For OSC, this involves spectral decomposition followed by matrix multiplication of $n \times K^2$ and $K^2 \times n$ matrices, so the complexity is $O(n^3 + n^2 K^2)$. 
For SSC-ASE, we again start with spectral decomposition, followed by $n$ LASSO problems with a design matrix of size $K^2 \times (n - 1)$, so the complexity is $O(n^3 + n^2 K^4 + n K^6)$ \cite{10.1214/009053604000000067}. 
SSC-A involves solving $n$ LASSO regression problems each with design matrices of size $n \times (n - 1)$, so the complexity here is $O(n^4)$. 
When $n \gg K^2$, we can see that OSC and SSC-ASE are both $O(n^3)$. 
In practice, we have found that $K$ does not affect runtimes too much. 
}

### 2022-01-11 referee-response.tex

  This has been added to the paper.
  }
\item
  p10. I didn't understand the exposition of the SSC algorithm.\\
  We have added more detail to our description of the SSC algorithm. 
  In particular, we talk a bit about the intuition/motivation behind the algorithm. 
  This isn't added to the description, but $c$ and $c_i$ are column vectors of length $n$ and $c^{(i)}$ and $c_i^{(j)}$ are the $i^{th}$ and $j^{th}$ entries of vectors $c$ and $c_i$ respectively. 
  \textcolor{blue}{The description of SSC now reads:
  The SSC algorithm can be described as follows: 
  Given \(X \in \mathbb{R}^{n \times d}\) with vectors
  \(x_i^\top \in \mathbb{R}^d\) as rows of \(X\), the optimization problem
  \(c_i = \arg\min_{c} \|c\|_1\) subject to \(x_i = X^\top c\) and
  \(c^{(i)} = 0\), where $c^{(i)}$ is the $i^{th}$ entry of $c$, is solved for each \(i = [n]\). 
  The solutions are collected collected into matrix
  \(C = \bigl[ c_1 \mid \cdots \mid c_n \bigr]^\top\) to
  construct an affinity matrix \(B = |C| + |C^\top|\). If each \(x_i\) lies
  exactly on one of \(K\) subspaces, \(B\) describes an undirected graph
  consisting of {\em at least} \(K\) disjoint subgraphs, i.e., \(B_{ij} = 0\) if \(x_i, x_j\) lie on different subspaces. 
  The intuition here is that points that lie on the same subspace can be described as linear combinations of each other. 
  Thus, for each $c_i$, $c_i^{(j)}$ is zero if $x_i$ and $x_j$ belong to different subspaces and may be nonzero if they do. 
  If \(X\) instead represents points near \(K\) subspaces with some noise, 
  then this property may only hold approximately and thus 
  a final graph partitioning step may be required 
  (e.g., edge thresholding or spectral clustering).

@@

  \textcolor{blue}{
  Regarding computational complexity, because both OSC and SSC-ASE may have an arbitrary final processing step, we will focus on the complexity in constructing the affinity matrix $B$. 
  For OSC, this involves spectral decomposition followed by matrix multiplication of $n \times K^2$ and $K^2 \times n$ matrices, so the complexity is $O(n^3 + n^2 K^2)$. 
  For SSC-ASE, we again start with spectral decomposition, followed by $n$ LASSO problems with a design matrix of size $K^2 \times (n - 1)$, so the complexity is $O(n^3 + n^2 K^4 + n K^6)$ \cite{10.1214/009053604000000067}. 

### 2022-01-11 summary_minh_edited.tex
more responses to referee comments

\citet{noroozi2019estimation} used Sparse Subspace Clustering (SSC) 
(\cite{5206547}) for community detection in the PABM. 
The SSC algorithm can be described as follows: 
Given \(X \in \mathbb{R}^{n \times d}\) with vectors
\(x_i^\top \in \mathbb{R}^d\) as rows of \(X\), the optimization problem
\(c_i = \argmin_{c} \|c\|_1\) subject to \(x_i = X^\top c\) and
\(c^{(i)} = 0\), where $c^{(i)}$ is the $i^{th}$ entry of $c$, is solved for each \(i = 1, ..., n\). The solutions are
collected collected into matrix
\(C = \bigl[ c_1 \mid \cdots \mid c_n \bigr]^\top\) to
construct an affinity matrix \(B = |C| + |C^\top|\). If each \(x_i\) lies
exactly on one of \(K\) subspaces, \(B\) describes an undirected graph
consisting of {\em at least} \(K\) disjoint subgraphs, i.e., \(B_{ij} = 0\) if
\(x_i, x_j\) lie on different subspaces. If \(X\) instead represents
points near \(K\) subspaces with some noise, a final graph partitioning
step may be required (e.g., edge thresholding or spectral clustering).

In practice, due to presence of noise, SSC is often done by solving the LASSO problems
\begin{equation} \label{eq:ssc}
c_i = \argmin_c \frac{1}{2} \|x_i - X_{-i}^\top c\|^2_2 + \vartheta \|c\|_1
\end{equation}
for some sparsity parameter \(\vartheta > 0\). The \(c_i\) vectors are
then collected into \(C\) and \(B\) as before.

>>>>

\citet{noroozi2019estimation} used Sparse Subspace Clustering (SSC) 
(\cite{5206547}) for community detection in the PABM. 
The SSC algorithm can be described as follows: 
Given \(X \in \mathbb{R}^{n \times d}\) with vectors
\(x_i^\top \in \mathbb{R}^d\) as rows of \(X\), the optimization problem
\(c_i = \argmin_{c} \|c\|_1\) subject to \(x_i = X^\top c\) and
\(c^{(i)} = 0\), where $c^{(i)}$ is the $i^{th}$ entry of $c$, is solved for each \(i = [n]\). 
The solutions are collected collected into matrix
\(C = \bigl[ c_1 \mid \cdots \mid c_n \bigr]^\top\) to
construct an affinity matrix \(B = |C| + |C^\top|\). If each \(x_i\) lies
exactly on one of \(K\) subspaces, \(B\) describes an undirected graph
consisting of {\em at least} \(K\) disjoint subgraphs, i.e., \(B_{ij} = 0\) if \(x_i, x_j\) lie on different subspaces. 
The intuition here is that points that lie on the same subspace can be described as linear combinations of each other. 
Thus, for each $c_i$, $c_i^{(j)}$ is zero if $x_i$ and $x_j$ belong to different subspaces and may be nonzero if they do. 
If \(X\) instead represents points near \(K\) subspaces with some noise, 
then this property may only hold approximately and thus 
a final graph partitioning step may be required 
(e.g., edge thresholding or spectral clustering).

In practice, due to presence of noise, SSC is often done by solving the LASSO problems

\begin{equation} \label{eq:ssc}
c_i = \argmin_c \frac{1}{2} \|x_i - X_{-i}^\top c\|^2_2 + \vartheta \|c\|_1
\end{equation}

for some sparsity parameter \(\vartheta > 0\). 
The \(c_i\) vectors are then collected into \(C\) and \(B\) as before.
